---
title: "How to analyze the accuracy and stability of a network using PRONA"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{introduction}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  markdown: 
    wrap: 72
---

This vignette gives an overview of how to perform item stability
analysis, nonparametric bootstrapping, casedrop bootstrapping, and
nodedrop bootstrapping on a GGM network generated with PRONA. For
running these analyses, PRONA largely serves as a wrapper for the
[EGANet](https://github.com/hfgolino/EGAnet) and
[bootnet](https://github.com/SachaEpskamp/bootnet) packages. For more
information on these analyses and how to interpret their results, please
see [Christensen & Golino, Psych,
2021](https://doi.org/10.3390/psych3030032) and [Epskamp, Borsboom, &
Fried, Behavioral Research Methods,
2018](https://pubmed.ncbi.nlm.nih.gov/28342071/).

### 1. Load PRONA, download data, and construct a GGM

For this example, we will use the data from the introduction vignette to
construct a GGM. Please refer to the introduction vignette for more
details on where this data is from and how the network is constructed.

```{r}
library(PRONA)
reduced_df <- read.csv('../../PRONA_additional_files/test_scripts/output/reduced_data.csv', stringsAsFactors = FALSE)
ptsd_network <- construct_ggm(reduced_df, normal = FALSE)
```

```{r, fig.height=6, fig.width=9}
plot_ggm(ptsd_network)
```

### 2. Perform Item Stability Analysis

To perform item stability analysis, we use the run_bootEGA function.
This function has 3 parameters:

-   df: A dataframe of symptom severity/frequency data

-   normal: Whether to consider all variables normally distributed. If
    false, conducts a non-paranormal transformation (Default: FALSE)

-   iter: Number of permutations to run for bootstrapping analysis
    (Default: 10,000)

```{r, results='hide'}
# Replace file path with wherever you want to save the results of bootEGA
boot.ega.file <- "../../PRONA_additional_files/test_scripts/output/boot.ega.RDS"
if (!file.exists(boot.ega.file)){
  boot.ega <- run_bootEGA(reduced_df)
  saveRDS(boot.ega, boot.ega.file)}
boot.ega <- readRDS(boot.ega.file)
```

Let's plot the median network structure from bootEGA.

```{r, fig.width=7, fig.height=5}
plot_bootEGA(boot.ega)
```

Next, we calculate item and dimension stability and plot the results

```{r, results='hide', fig.width=7, fig.height=5}
dimStab <- calculate_dimStab(boot.ega)
```

```{r, fig.width=7, fig.height=5}
plot_item_stability(dimStab)
```

### 3. Perform nonparametric bootstrapping to analyze edge weight accuracy

To perform nonparametric bootstrapping, we use the nonparam_boot
function, which has five parameters:

-   df: A dataframe of symptom severity/frequency data

-   normal: Whether to consider all variables normally distributed. If
    false, conducts a non-paranormal transformation (Default: FALSE)

-   nBoots: Number of permutations to run for bootstrapping analysis
    (Default: 2,500)

-   statistics: Vector indicating which statistics to store. See bootnet
    for options. (Default:
    c('edge','strength','closeness','betweenness')

-   communities: If you are running bootnet on bridge centrality
    measures, use this parameter to set community labels (set this to
    ega\$wc, where ega is the output from the construct_ggm function).

```{r, results='hide'}
nonparam.boot.file <- "../../PRONA_additional_files/test_scripts/output/nonparam.boot.RDS"
if (!file.exists(nonparam.boot.file)){
  nonparam.boot <- nonparam_boot(reduced_df, normal = FALSE, statistics = c('edge','strength','closeness','betweenness','bridgeStrength','bridgeCloseness','bridgeBetweenness'), communities = ptsd_network$wc)
  saveRDS(nonparam.boot, nonparam.boot.file)}
nonparam.boot <- readRDS(nonparam.boot.file)
```

Let's plot and summarize the results.

```{r, fig.width=7, fig.height=5}
library(bootnet)
plot_nonparam_boot(nonparam.boot)
```

```{r}
summarize_nonparam_boot(nonparam.boot)
```

### 4. Perform casedrop and nodedrop bootstrapping to analyze edge weight and centrality stability

To perform casedrop bootstrapping, we use the casedrop_boot function,
which has the same five parameters as the nonparam_boot function.

```{r, results='hide'}
casedrop.boot.file <- "../../PRONA_additional_files/test_scripts/output/casedrop.boot.RDS"
if (!file.exists(casedrop.boot.file)){
  casedrop.boot <- casedrop_boot(reduced_df, normal = FALSE, statistics = c('edge','strength','closeness','betweenness','bridgeStrength','bridgeCloseness','bridgeBetweenness'), communities = ptsd_network$wc)
  saveRDS(casedrop.boot, casedrop.boot.file)}
casedrop.boot <- readRDS(casedrop.boot.file)
```

Plot the casedrop bootstrapping results for edge weight, centralities,
and bridge centralities

```{r, fig.width=7, fig.height=5}
plot_casedrop_boot(casedrop.boot, type = "edge")
```

```{r, fig.width=7, fig.height=5}
plot_casedrop_boot(casedrop.boot, type = "centralities")
```

```{r, fig.width=7, fig.height=5}
plot_casedrop_boot(casedrop.boot, type = "bridge centralities")
```

Next, we perform correlation stability analysis on the casedrop
bootstrapping results.

```{r}
cor_stability_analysis(casedrop.boot)
```

Finally, let's perform nodedrop bootstrapping and plot the results.

```{r, results='hide'}
nodedrop.boot.file <- "../../PRONA_additional_files/test_scripts/output/nodedrop.boot.RDS"
if (!file.exists(nodedrop.boot.file)){
  nodedrop.boot <- nodedrop_boot(reduced_df, normal = FALSE, statistics = c('edge','strength','closeness','betweenness'))
  saveRDS(nodedrop.boot, nodedrop.boot.file)}
nodedrop.boot <- readRDS(nodedrop.boot.file)
```

```{r, fig.width=7, fig.height=5}
plot_nodedrop_boot(nodedrop.boot, type = 'edge')
```

```{r, fig.width=7, fig.height=5}
plot_nodedrop_boot(nodedrop.boot, type = 'centralities')
```
