---
title: "Perform concordance network-based clustering using PRONA"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{introduction}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  markdown: 
    wrap: 72
---

This vignette gives an overview of how to perform second-order,
concordance network-based unsupervised clustering to identify
communities of patients with unique symptom network patterns. For more
information on this clustering method and examples of its applications,
please see [Henry et al., Plos One,
2018](https://doi.org/10.1371/journal.pone.0191981) and [Bergsneider et
al., Neuro-Oncology Advances,
2022](https://doi.org/10.1093/noajnl/vdac188).

### 1. Load PRONA, download data, and construct a GGM

For this example, we will use the data from the introduction vignette to
construct a GGM. Please refer to the introduction vignette for more
details on where this data is from and how the network is constructed.

```{r}
library(PRONA)
reduced_df <- read.csv('../../PRONA_additional_files/test_scripts/output/reduced_data.csv', stringsAsFactors = FALSE)
ptsd_network <- construct_ggm(reduced_df, normal = FALSE)
```

```{r, fig.width=9, fig.height=6}
plot_ggm(ptsd_network)
```

### 2. Get communities of patients with unique symptom severity/frequency patterns

To run second-order unsupervised clustering, we use the get_communities
function, which has 6 parameters:

-   data: Dataframe of symptom severities (formatted as required by
    PRONA)

-   nrows: Number of rows in the dataframe (Defaults to the number of
    rows in data, you should not change this parameter)

-   ncols: Number of columns in the dataframe (Defaults to the number of
    columns in data, you should not change this parameter)

-   detectAlgo: The community detection algorithm to use. Options
    include FG (fast greedy), IM (infomap), LP (label propagation), LE
    (leading eigen), LV (louvain), or WT (walktrap). (Default: WT)

-   simil_measure: The similarity measure to use, either 'Euclidean' or
    'ARI' (Adjusted Rand Index) (Default: ARI)

-   simplify_graphs: Boolean that indicates whether to simplify the
    graph by removing multi-edges and loops (Default: TRUE)

It returns a new dataframe that is a copy of the original symptom
severities dataframe, but with a new column representing which community
each patient belongs to.

```{r}
community_df <- get_communities(reduced_df)
```

Lets see how many patients are in each community

```{r}
table(community_df$community)
```

Next, lets plot a heatmap of symptom severities for each community using
the plot_community_heatmap function, which has three parameters:

-   df: The output of the get_communities function

-   cluster_rows: Boolean representing whether or not to cluster the
    rows (symptoms) via heirarchical clustering (Default: TRUE)

-   symptom_order: A vector of symptom names indicating the order to
    plot symptoms on the y-axis. This order will only show up if
    cluster_rows = FALSE. The default order if cluster_rows = FALSE is
    the order of column names in the dataframe (minus the ID and
    community columns)

Here, "PC" stands for "Patient Community":

```{r, fig.width=7, fig.height=5}
plot_community_heatmap(community_df)
```

We can also visualize the symptom severities of each community using
line plots. The community_line_plot function has two parameters:

-   data: output of get_communities

-   communities: ector specifying which communities to plot. If you
    include 0 in the vector, it will plot the severity for all patients
    combined. (Default: c(0))

```{r, fig.width=7, fig.height=5}
community_line_plot(community_df, c(0,1,2))
```

We can also get a summary of each symptom in each community using the
get_community_summary function, which returns a dataframe with the mean,
median, upper bound, and lower bound of each symptom in each community.

```{r}
get_community_summary(community_df)
```

### 3. Construct networks for individual communities

We can construct GGM networks and conduct statistical assessment for
individual communities using the same pipeline as described in the
introduction and statistical assessment vignettes.

```{r}
# Get symptom data for individual communities
library(dplyr)
comm1_df <- community_df[community_df$community==1,] %>% select(-community)
comm2_df <- community_df[community_df$community==2,] %>% select(-community)
```

```{r}
comm1_ega <- construct_ggm(comm1_df, normal = FALSE)
```

```{r, fig.width=7, fig.height=5}
plot_ggm(comm1_ega)
```

```{r}
comm2_ega <- construct_ggm(comm2_df, normal = FALSE)
```

```{r, fig.width=7, fig.height=5}
plot_ggm(comm2_ega)
```

Because of its relatively low number of patients and low overall symptom
severity, the network for PC1 is sparse.

### 4. Perform Network Comparison Test (NCT)

Finally, we can statistically assess the differences between the two
networks using the Network Comparison Test. For details on how the
Network Comparison Test works, please see [Borkulo et al., Psychol
Methods, 2022](https://pubmed.ncbi.nlm.nih.gov/35404628/). The functions
in PRONA serve as a wrapper for the functions in the
[NetworkComparisonTest
package](https://github.com/cvborkulo/NetworkComparisonTest).

The run_NCT function takes 8 parameters:

-   df1: First dataframe of symptom severity/frequency data

-   df2: Second dataframe of symptom severity/frequency data

-   normal: Boolean. Whether to consider all variables normally
    distrubted. If false, conducts a non-paranormal transformation
    (Default: FALSE)

-   it: Number of bootstrapping iterations to run (Default: 2500)

-   p.adjust.methods: Character. Can be one of "holm", "hochberg",
    "hommel", "bonferroni", "BH", "BY", "fdr", or "none". To control (or
    not) for testing of multiple edges. Defaults to "none".

-   test.edges: Boolean. Whether to test differences in individual edge
    weights. (Default: TRUE)

-   test.centrality: Boolean. Whether to test differences in centrality
    measures (Default: TRUE)

-   centrality: Vector of which centrality measures to test (Default:
    c('closeness','betweenness','strength','expectedInfluence'))

```{r, results='hide'}
# Replace file path with wherever you want to save the results of NCT
nct.file <- "../../PRONA_additional_files/test_scripts/output/nct.RDS"
if (!file.exists(nct.file)){
  nct.results <- run_NCT(comm1_df, comm2_df)
  saveRDS(nct.results, nct.file)}
nct.results <- readRDS(nct.file)
```

We can plot the results of the NCT comparing global network strength and
overall network structure.

```{r, fig.width=7, fig.height=5}
library(NetworkComparisonTest)
plot_NCT(nct.results, what = "strength")
```

```{r, fig.width=7, fig.height=5}
plot_NCT(nct.results, what = "network")
```

We can also directly get the p-values from comparing each edge weight
and centrality measure

```{r}
get_NCT_pvalues(nct.results, what = "edges")
```

```{r}
get_NCT_pvalues(nct.results, what = "centralities")
```
