---
title: 'P1: Latent Variable Reduction'
author: "Brandon Bergsneider"
date: "2/11/2022"
output: html_document
---

### Analyzing the MDASI-BT data for 1128 PBT patients, this script identifies redundant variables
### and consolidates them into latent variables using the Unique Variable Analysis method described in Christensen, Garrido, & Golino (PsyArXiv)

```{r}
# Load packages
library('EGAnet')
library("dplyr")
library("tidyr")
library("qgraph")
library("ggplot2")
library("reshape2")
library(readr)

orange="#ff7f00"
green="#4caf4a"
red="#e4211c"
blue="#387db8"
```

```{r, results=FALSE, message=FALSE, warning=FALSE}
# Load data
symptom_data <- read_csv('data/Merged_SymptomSeverity_Upload_Final.csv')
```

```{r}
# Create dataframe for symptom data from all patients
all_symptom_data <- dplyr::select(symptom_data, -Study, -ID)
```

### Calculate how many patients report at least one symptom, how many patients report at least one severe symptom

```{r}
# Number of patients with at least one symptom
sum(rowSums(all_symptom_data)==0)
```

```{r}
# Number of patients with at least one severe symptom (>= 5)
count(filter_all(all_symptom_data, any_vars(. >= 5)))
```

```{r}
# Number of patients with concurrent symptoms
tmp <- all_symptom_data
tmp[tmp>0] <- 1
table(rowSums(tmp))
```


### Construct network using all symptoms

```{r}
# Function for constructing a GGM regularized using EBICglasso method
# All variables considered continuous
# If covariance matrix is nonpositive definite, matrix is forced to be nonpositive definite using the nearPD function
# Input variables:
# - data: dataframe of symptom severity data to analyze
# - npn: boolean representing whether or not to consider data as nonparanormal (not continuous)
construct_network <- function(data, npn=FALSE) {
  cor_matrix <- cor_auto(data, detectOrdinal=FALSE, forcePD=TRUE, npn.SKEPTIC=npn)
  GLASSO_matrix <- EBICglasso(cor_matrix, nrow(data), 0.5)
  network <- qgraph(GLASSO_matrix, layout = "spring")
  return(network)
}
```

```{r}
# Construct a GGM from the original symptom data to see which variables are highly correlated
construct_network(all_symptom_data)
```

### From visual inspection, the symptom pairs vmt-nas, sad-dst, wkn-nmb, and ftg-drw all appear highly correlated and may be redundant

### Lets look at the severity and occurrence distributions of each variable to determine which variables may be redundant
### Although not established in network analysis, Exporatory Factor Analysis (EFA) methods commonly remove symptoms present
### in <20% or >80% of patients in order to have sufficient variation and covariation in the data

```{r}
tmp_all_symptom_data <- all_symptom_data
tmp_all_symptom_data['comm'] <- 1
long_all_symptom_data <- melt(tmp_all_symptom_data, id='comm')
```

```{r, fig.width=3.5, fig.height=4.5}
severity_plot <- ggplot(long_all_symptom_data, aes(x=reorder(variable,value), y=value)) + geom_boxplot(fill=green)  + ylab('severity\n') + xlab('') + theme_minimal() + coord_flip() + theme(axis.text = element_text(size = 12), axis.title = element_text(size = 12)) 
severity_plot
#ggsave('images/severity_plot.png', plot = severity_plot, width = 8, height = 5)
```

```{r}
tmp_all_symptom_data <- all_symptom_data
frequency_all <- round(colSums( tmp_all_symptom_data != 0)/nrow(tmp_all_symptom_data)*100,1)
frequency_df <- data.frame(symptom=names(frequency_all), frequency=unname(frequency_all))
frequency_df
```

```{r, fig.width=3.5, fig.height=4.5}
frequency_plot <- ggplot(data=frequency_df, aes(x=reorder(symptom,frequency),y=frequency)) + geom_bar(stat="identity", color='black', fill=red, alpha=0.75, width=0.8) + ylab('% occurrence\n(mild-severe)') + xlab('') + ylim(0,100) + geom_text(aes(label=sprintf("%.1f",frequency)),hjust=-0.3, size=4) + theme_minimal() + coord_flip() + theme(axis.text = element_text(size = 12), axis.title = element_text(size = 12)) 

frequency_plot
#ggsave('images/mild_frequency_plot.png', plot = frequency_plot, width = 8, height = 5)
```

```{r}
tmp_all_symptom_data <- all_symptom_data
frequency_severe <- round(colSums( tmp_all_symptom_data > 4)/nrow(tmp_all_symptom_data)*100,1)
severe_frequency_df <- data.frame(symptom=names(frequency_severe), frequency= unname(frequency_severe))
severe_frequency_df
```

```{r, fig.width=3.5, fig.height=4.5}
severe_frequency_plot <- ggplot(data=severe_frequency_df, aes(x=reorder(symptom,frequency),y=frequency)) + geom_bar(stat="identity", color='black', fill=blue, alpha=0.75, width=0.8) + ylab('% occurrence\n(moderate-severe)') + xlab('') + ylim(0,100) + geom_text(aes(label=sprintf("%.1f",frequency)),hjust=-0.3, size=4) + theme_minimal() + coord_flip() + theme(axis.text = element_text(size = 12), axis.title = element_text(size = 12))
severe_frequency_plot
#ggsave('images/severe_frequency_plot.png', plot = severe_frequency_plot, width = 8, height = 5)
```

### Based on the symptom severity and occurrence distributions, it seems that vomiting (present in 9% of pts) and fatigue (present in 79% of pts)
### might be good variables to consolidate

### Next, perform Unique Variable Analysis (UVA) to consolidate redundant variables

```{r}
if (!file.exists("data/uvaResults.RDS")){
  # Perform manual UVA (latent variable)
  data.manual <- UVA(data = all_symptom_data, auto = FALSE)
  saveRDS(data.manual, "data/uvaResults.RDS")
}
data.manual <- readRDS("data/uvaResults.RDS")
```

### After UVA, redundant variables with both topological and conceptual overlap were combined into latent variables
### Five redundant pairs were identified by UVA: vmt-nas, sad-dst, wkn-nmb, ftg-drw, and und-spk
### However, only vmt-nas and ftg-drw were combined into latent variables because they had substantial conceptual overlap
### Furthermore, as noted above, vomiting and fatigue are good variables to consolidate because
### of their low and high occurrence rates, respectively

```{r}
# Check wTO between all pairs of variables
data.manual$redundancy$descriptives
```

```{r, fig.width=5, fig.height=4.5}
# Check wTO between all pairs of variables
wTO_df <- data.frame(data.manual$redundancy$descriptives$centralTendency)
wTO_df <- cbind(pair = rownames(wTO_df), wTO_df)
rownames(wTO_df) <- 1:nrow(wTO_df)
head(wTO_df,10)

wTO_plot <- ggplot(data=head(wTO_df,20), aes(x=reorder(pair,wTO),y=wTO)) + geom_bar(stat="identity", color='black', fill=orange, alpha=0.75, width=0.8) + ylab('wTO\n') + xlab('') + ylim(0,0.5) + geom_text(aes(label=sprintf("%.3f",wTO)),hjust=-0.3, size=4) + theme_minimal() + coord_flip() + theme(axis.text = element_text(size = 12), axis.title = element_text(size = 12))
wTO_plot
#ggsave('images/wTO_plot.png', plot = wTO_plot, width = 8, height = 5)
```


```{r}
# Save reduced data in a new dataframe
reduced_all_symptom_data <- data.manual$reduced$data

# Latent variable reduction yields variable values that are not between 0 and 10
# In order to keep latent variable values consistent with other variables, latent variable
# values were normalized to be between 0 and 10
for(lv in list('fatigue-drowsy','nausea-vomiting')) {
  reduced_all_symptom_data[,lv] <- (reduced_all_symptom_data[,lv]-min(reduced_all_symptom_data[,lv]))/(max(reduced_all_symptom_data[,lv])-min(reduced_all_symptom_data[,lv]))*10
}
```

```{r, fig.width=3, fig.height=4.5}
library(ggridges)
reduced_all_symptom_data_long <- reduced_all_symptom_data %>% data.frame() %>% pivot_longer(everything(), names_to = "symptom", values_to="severity") %>% rowwise() %>% mutate(symptom=gsub("\\.","-", symptom))

density_plot <- ggplot(reduced_all_symptom_data_long, aes(x=severity, y=symptom)) + geom_density_ridges(fill='#4caf4a')  + xlab('severity\ndistribution') + ylab('') + theme_minimal() + theme(axis.text = element_text(size = 12), axis.title = element_text(size = 12)) +
  scale_y_discrete(expand = c(0, 0)) +
  scale_x_continuous(expand = c(0, 0)) + 
  coord_cartesian(clip = "off") 
density_plot

apply(reduced_all_symptom_data, 2, shapiro.test)

```

```{r}
# Check that latent variable normalization does not affect structure of the network
n1 <- construct_network(reduced_all_symptom_data)
n2 <- construct_network(data.manual$reduced$data)
```

```{r}
# Save reduced symptom dataframe as a csv file
#write.csv(reduced_all_symptom_data, '~/OneDrive-NIH/Unsupervised_Clustering/data/p1_reduced_all_symptom_data.csv', row.names = FALSE)
```

