---
title: 'P2: All Patient Network'
author: "Brandon Bergsneider"
date: "2/11/2022"
output: html_document
---

### This script constructs a GGM for all 1128 PBT patients, identifies symptom clusters

### within the network, and analyzes the accuracy and stability of the network

```{r}
# Load packages
library("dplyr")
library("tidyr")
library("qgraph")
library("EGAnet")
library("bootnet")
library("networktools")
library("ggpubr")


orange="#ff7f00"
green="#4caf4a"
red="#e4211c"
blue="#387db8"

centrality_colors <- c("Strength"=red, "Closeness"=blue, "Betweenness"=orange)
bridge_centrality_colors <- centrality_colors
names(bridge_centrality_colors) <- sapply(names(bridge_centrality_colors), function(x) paste0("Bridge\n",x))
```

```{r, results=FALSE, message=FALSE, warning=FALSE}
# Load data
symptom_data <- read.csv('data/p1_reduced_all_symptom_data.csv', stringsAsFactors = FALSE)
```

### Start by assessing the normality of each variable using Shapiro-Wilk's Test

```{r}
apply(symptom_data, 2, shapiro.test)
```

### Based on the Shapiro-Wilks Normality Test, none of the variables in our dataset are normally distributed

### Thus, when constructing graphs we will have to use the non-paranormal transformation

### Now, lets construct a GGM of all patients

```{r}
# Calculate the correlation matrix
# Assume that all variables are continuous and non-normal
cor_matrix <- cor_auto(symptom_data, detectOrdinal = FALSE, npn.SKEPTIC = TRUE, forcePD = TRUE)
```

```{r}
# Construct an EBIC-GLASSO regularized GGM using the EGA function from EGAnet
# EGA also identifies symptom clusters using the walktrap algorithm
rownames(cor_matrix) <- gsub("\\.","-", rownames(cor_matrix))
colnames(cor_matrix) <- gsub("\\.","-", colnames(cor_matrix))

ega <- EGA(cor_matrix, n = nrow(symptom_data), plot.EGA = FALSE)
```

```{r,fig.width=6, fig.height=3.5}
# Plot the EGA
plot.ega <- plot(ega, plot.args = list(node.size = 8, label.size = 3.5, color.palette = c(orange, red, green, blue), legend.names=c("cognitive", "physical", "focal neurologic", "affective")))

#ggsave('images/all_pt_EGA_with_clusters.png', plot = plot.ega, width = 8, height = 5)
```

```{r}
# Plot the EGA
plot.ega <- plot(ega, plot.args = list(node.size = 8, label.size = 3.5, color.palette = c(orange, red, green, blue), legend.names=c("cognitive", "physical", "focal neurologic", "affective")))
```

```{r, eval=FALSE}
plot.ega <- plot(ega, plot.args = list(node.size = 8, label.size = 6))
plot.ega
#ggsave('poster_images/all_pt_EGA_with_clusters.png', plot = plot.ega, width = 8, height = 5)
```

```{r, fig.width=6, fig.height=3.5}
# Plot the EGA without community identification
plot.ega <- plot(ega, plot.args = list(node.size = 8, label.size =3.5, color.palette = c(green, green, green, green), legend.names=c("cognitive", "physical", "focal neurologic", "affective")))
#ggsave('images/all_pt_EGA.png', plot = plot.ega, width = 8, height = 5)
```

```{r}
# Print number of edges in the graph and a dataframe of edge weights
m <- ega$network
print(paste0('# egdes = ',length(ega$network[ega$network != 0][upper.tri(m)]),'/400'))
DT::datatable(data.frame(row=rownames(m)[row(m)[upper.tri(m)]], 
           col=colnames(m)[col(m)[upper.tri(m)]], 
           weight=m[upper.tri(m)]))
```

### Calculate the centrality measures of the network

```{r, fig.width=5, fig.height=4.5}
centralities <- centralityTable(ega$network) %>% filter(measure %in% c("Strength", "Closeness", "Betweenness")) %>% select(node, measure, value)
orderbystrength <- centralities%>%filter(measure=="Strength") %>% arrange(value) %>% select(node)
centralities <- centralities %>% group_by(measure) %>% mutate(node = factor(node, levels=orderbystrength$node), measure=factor(measure, levels=c("Strength", "Closeness", "Betweenness"))) %>% arrange(node)
p <- ggplot(centralities %>% group_by(measure), aes(x=value, y=node, group=measure, color=measure)) + 
  geom_point(size=2) + geom_path(size=0.7) +  
  scale_color_manual(values=centrality_colors) +
  facet_wrap(~measure, ncol=3) + theme_minimal() +
  xlab("") + 
  ylab("") + 
  theme(axis.text.y =element_text(size=12),
        axis.text.x=element_text(size=12),
        strip.text=element_text(size=12), 
        legend.position = "none")
p

```

```{r, fig.width=4.5, fig.height=4.5}
p <- ggplot(centralities %>% group_by(measure, node), aes(x=value, y=node, group=measure, color=measure)) + 
  geom_point(size=2) + geom_path(size=0.7) +  
  scale_color_manual(values=centrality_colors) +
  #facet_wrap(~measure, ncol=3) + 
  theme_minimal() +
  xlab("") + 
  ylab("") + 
  theme(axis.text.y =element_text(size=12),
        axis.text.x =element_text(size=12),
        strip.text=element_text(size=12), 
        legend.text = element_text(size=12), legend.title=element_blank())
p
```

```{r, fig.width=2.5, fig.height=4.5, eval=FALSE}
sapply(c("Strength", "Closeness", "Betweenness"), function(x){
  centralities <- centralityTable(ega$network) %>% filter(measure %in% c(x)) %>% select(node, measure, value)
  orderbyvalue <- centralities%>%arrange(value) %>% select(node)
  cc <- unname(centrality_colors[x])
  centralities <- centralities %>% mutate(node = factor(node, levels=orderbyvalue$node)) 
  p <- ggplot(centralities %>% data.frame, aes(x=value, y=node, group=measure)) + 
    geom_point(color=cc) + geom_line(color=cc) + theme_minimal() +
    xlab(x) + 
    ylab("") + 
    xlim(-4,4) +
    theme(axis.text.y =element_text(size=12),
          axis.text.x = element_text(size=12),
          strip.text=element_text(size=12))
  print(p)
})

# Calculate and plot the centrality measures of the network
#all_pt_centrality_plot <- centralityPlot(ega$network, scale = "z-scores", include = c('Strength','Closeness','Betweenness'), orderBy = 'Strength') + theme(text = element_text(size = 22.5))
#ggsave('images/all_pt_centrality_plot.png', plot = all_pt_centrality_plot, width = 8, height = 5)
```

```{r, eval=FALSE}
# Calculate and plot the centrality measures of the network
all_pt_centrality_plot <- centralityPlot(ega$network, scale = "z-scores", include = c('Strength','Closeness','Betweenness'), orderBy = 'Strength') + theme(text = element_text(size = 22.5))
#ggsave('poster_images/all_pt_centrality_plot.png', plot = all_pt_centrality_plot, width = 8, height = 5)
```

```{r}
centrality_stats <- centrality_auto(ega$network)
centrality_table <- centrality_stats$node.centrality
DT::datatable(centrality_table)
```

### Calculate bridge centrality measures

```{r}
# Calculate and plot the centrality measures of the network
bridge_centralities <- bridge(ega$network, communities=ega$wc)
```

```{r, eval=FALSE}
plot(bridge_centralities, zscore = TRUE, include = c('Bridge Strength','Bridge Closeness','Bridge Betweenness'), color = FALSE, order = 'value')
```

```{r, eval=FALSE}
bridge_centrality_plot <- plot(bridge_centralities, zscore = TRUE, include = c('Bridge Strength','Bridge Closeness','Bridge Betweenness'), color = FALSE, order = 'value')
#ggsave('images/bridge_centrality_plot.png', plot = bridge_centrality_plot, width = 8, height = 5)
```

```{r}
bridge_centrality_table <- as.data.frame(list(bridge_centralities$'Bridge Strength', bridge_centralities$'Bridge Betweenness', bridge_centralities$'Bridge Closeness'), col.names = list('Bridge Strength','Bridge Betweenness','Bridge Closeness'))
DT::datatable(bridge_centrality_table)
```

```{r, fig.width=5, fig.height=4.5}
bridge_centrality_scaled <- apply(bridge_centrality_table, 2, scale) %>% as.data.frame()
bridge_centrality_scaled$node = rownames(bridge_centrality_table)
bridge_centrality_scaled <- bridge_centrality_scaled %>% pivot_longer(!node, names_to="measure", values_to = "value")
bridge_centrality_scaled <- bridge_centrality_scaled %>% rowwise() %>% mutate(measure=gsub("\\.", "\n", measure))

orderbystrength <- bridge_centrality_scaled%>%filter(measure=="Bridge\nStrength") %>% arrange(value) %>% select(node)
bridge_centrality_scaled <- bridge_centrality_scaled %>% group_by(measure) %>% mutate(node = factor(node, levels=orderbystrength$node), measure=factor(measure, levels=c("Bridge\nStrength", "Bridge\nCloseness", "Bridge\nBetweenness"))) %>% arrange(node)
p <- ggplot(bridge_centrality_scaled %>% group_by(measure), aes(x=value, y=node, group=measure, color=measure)) + 
  geom_point(size=2) + geom_path(size=0.7) +  
  scale_color_manual(values=bridge_centrality_colors) +
  facet_wrap(~measure, ncol=3) + theme_minimal() +
  xlab("") + 
  ylab("") + 
  theme(axis.text.y =element_text(size=12),
        axis.text.x=element_text(size=12),
        strip.text=element_text(size=12), 
        legend.position = "none")
p

```

### Analyze the stability of the network

### Conduct symptom cluster stability analysis using EGAnet

```{r}
if (!file.exists("data/boot.ega.RDS")){
  # Analyze stability of dimensions using parametric bootstrapping with 10,000 iterations
  boot.ega <- bootEGA(cor_matrix, n = nrow(symptom_data), iter = 10000, type = 'parametric')
  saveRDS(boot.ega, "data/boot.ega.RDS")}
boot.ega <- readRDS("data/boot.ega.RDS")
```

```{r, eval=FALSE}
# Plot median network structure
plot(boot.ega, plot.args = list(node.size = 8, label.size = 4))
```

![]()

```{r}
plot(boot.ega, plot.args = list(node.size = 8, label.size = 3.5, legend.names=c("cognitive", "physical", "affective", "focal neurolgic"), color.palette = c(blue,red,orange,green)))
```

```{r}
median_boot_ega <- plot(boot.ega, plot.args = list(node.size = 8, label.size = 4))
#ggsave('images/median_boot_ega.png', plot = median_boot_ega, width = 8, height = 5)
```

```{r}
# Print the proportion of times the number of dimensions was identified
boot.ega$frequency
```

```{r, fig.width=3, fig.height=4.5}
# Calculate both item and dimension stability
# Plots
dimStab <- dimensionStability(boot.ega)
```

```{r, fig.width=3.2, fig.height=4}
dimStab$item.stability$plot + theme_minimal() + scale_color_manual(values=c(red, blue, green, orange)) + 
  ylab("replication proportion") + 
  xlab("") + 
  theme(axis.text.y =element_text(size=11),
        axis.text.x=element_text(size=11),
        strip.text=element_text(size=11), 
        axis.title.y=element_text(size=11),
        legend.position = "none") 
```

```{r}
# Frequencies of symptom replication in each community
dimStab$item.stability$item.stability$all.dimensions
```

```{r}
# Node loadings for each community based on partial correlations
# Node loadings reflect each nodes contribution to the coherence of the communities
# They are interepreted in a similar way to exploratory factor analysis loadings, but as they
# are based on partial correlations, loadings >=0.2 indicate a medium association and >=0.3 strong association
dimStab$item.stability$mean.loadings
```

```{r}
typeof(dimStab$item.stability$mean.loadings)
```

```{r}
# Check the dimension stability: the proportion of times the original dimension is exactly replicated in across bootstrap samples
dimStab$dimension.stability$structural.consistency
# Check the average item stability in each empirical EGA dimension
dimStab$dimension.stability$average.item.stability
```

### Analyze edge weight accuracy using bootnet

```{r}
# Estimation function used for bootstrapping
# Calculates covariance matrix in same way as construct_network function
network_estimation_fun <- function(data) {
  cor_matrix <- cor_auto(data, detectOrdinal=FALSE, forcePD=TRUE, npn.SKEPTIC = TRUE)
  EBIC_matrix <- EBICglasso(cor_matrix, nrow(data), 0.5)
  return(EBIC_matrix)
}
```

```{r, results=FALSE, message=FALSE, warning=FALSE}
if (!file.exists("data/nonparam_boot.RDS")){
  # Conduct non-parametric bootstrapping to analyze edge weight accuracy
  nonparam_boot <- bootnet(symptom_data, nBoots=2500, statistics=c('edge','strength','closeness','betweenness','bridgeStrength','bridgeCloseness','bridgeBetweenness'), fun='network_estimation_fun', communities=ega$wc)
  saveRDS(nonparam_boot, "data/nonparam_boot.RDS")
}
nonparam_boot <- readRDS("data/nonparam_boot.RDS")
```

```{r, message=FALSE, warning=FALSE, fig.width=3, fig.height=4.5}
# Analyze edge weight accuracy
p <- plot(nonparam_boot, labels = FALSE, order = "sample", sampleColor=red, bootColor=blue, meanColor=blue, meanlwd=0.4, bootAlpha=0.2, bootlwd=0.4, areaAlpha=0.5)
pp <- p + theme_minimal() + theme(legend.position = "top", axis.text.y = element_blank(), 
                                  legend.text = element_text(size=12),
                                  axis.text.x=element_text(size=12),
                                  strip.text=element_text(size=12), 
                                  panel.grid.major.y = element_blank(),
                                  panel.grid.minor.y = element_blank())
print(pp)

summary_table <- summary(nonparam_boot, statistics = "edge")
DT::datatable(summary_table)
# Testing for significant differences in edge weights, without Bonferroni correction, for reasons details in Epskamp, Borsboom, & Fried 2017
p <- plot(nonparam_boot, "edge", plot="difference", onlyNonZero=TRUE, order="sample", labels=FALSE)

```

```{r}
nonparam_boot_plot <- plot(nonparam_boot, labels = FALSE, order = "sample") + theme(axis.text = element_text(size = 15), legend.text = element_text(size = 15))
#ggsave('images/nonparam_boot_plot.png', plot = nonparam_boot_plot, width = 8, height = 5)
```

```{r}
nonparam_boot_edge_comparison_plot <- plot(nonparam_boot, "edge", plot="difference", onlyNonZero=TRUE, order="sample", labels=FALSE) + theme(text = element_text(size = 20))
#ggsave('images/nonparam_boot_edge_comparison_plot.png', plot = nonparam_boot_edge_comparison_plot, width = 8, height = 5)
nonparam_boot_edge_comparison_plot
```

```{r}
summary_table['CIwidth'] <- summary_table['CIupper'] - summary_table['CIlower']
print(paste0('Median CI width for non-zero edges = ',median(summary_table[summary_table['sample']!=0,][['CIwidth']])))
print(paste0('Range of CI width for non-zero edges = ',min(summary_table[summary_table['sample']!=0,][['CIwidth']]),'-',max(summary_table[summary_table['sample']!=0,][['CIwidth']])))
```

```{r}
# Conduct casedrop bootstrapping
if (!file.exists("data/casedrop_boot.RDS")){
  casedrop_boot <- bootnet(symptom_data, nBoots=2500, type="case", statistics=c('edge','strength','closeness','betweenness','bridgeStrength','bridgeCloseness','bridgeBetweenness'), fun='network_estimation_fun', communities=ega$wc)
  saveRDS(casedrop_boot, "data/casedrop_boot.RDS")
}
casedrop_boot <- readRDS("data/casedrop_boot.RDS")
```

```{r}
# Plot casedrop bootstrapping results for edge weights
plot(casedrop_boot, statistics=c('edge'))
corStability(casedrop_boot)
```

```{r, fig.width=3.6, fig.height=3.5}
casedrop_boot_edge_plot <- plot(casedrop_boot, statistics=c('edge')) + theme_minimal() + xlab("% sampled cases") + ylab("correlation with original") + theme(axis.text.y =element_text(size=20),axis.text.x=element_text(size=20),axis.title =element_text(size=20), legend.text = element_text(size=30), legend.position = "top")
casedrop_boot_edge_plot
#ggsave('images/casedrop_boot_edge_plot.png', plot = casedrop_boot_edge_plot, width = 8, height = 5)
```

### Analyze centrality measure stability

```{r, fig.width=3.6, fig.height=3.5}
# Plot casedrop bootstrapping results for centrality measures
tolower_centrality_colors <- centrality_colors
names(tolower_centrality_colors) <- tolower(names(tolower_centrality_colors))
tobridge_centrality_colors <- centrality_colors
names(tobridge_centrality_colors) <- sapply(names(tobridge_centrality_colors), function(x) paste0("bridge",x))

plot(casedrop_boot, statistics=c('strength','closeness','betweenness')) + scale_color_manual(values=tolower_centrality_colors, labels=names(centrality_colors)) + scale_fill_manual(values=tolower_centrality_colors, labels=names(centrality_colors)) + theme_minimal() + xlab("% sampled cases") + ylab("correlation with original") + theme(axis.text.y =element_text(size=11),axis.text.x=element_text(size=10), legend.position = "top") + guides(fill=guide_legend(title=""), color=guide_legend(title=""))

plot(casedrop_boot, statistics=c('bridgeStrength','bridgeCloseness','bridgeBetweenness')) + scale_color_manual(values=tobridge_centrality_colors, labels=names(bridge_centrality_colors)) + scale_fill_manual(values=tobridge_centrality_colors, labels=names(bridge_centrality_colors)) + theme_minimal() + xlab("% sampled cases") + ylab("correlation with original") + 
  theme(axis.text.y =element_text(size=11),axis.text.x=element_text(size=10), legend.position = "top") + guides(fill=guide_legend(title=""), color=guide_legend(title=""))
corStability(casedrop_boot)
```

```{r}
casedrop_boot_plot <- plot(casedrop_boot, statistics=c('strength','closeness','betweenness')) + theme(text = element_text(size = 15))
#ggsave('images/casedrop_boot_plot.png', plot = casedrop_boot_plot, width = 8, height = 5)
```

```{r}
#Test for significant centrality differences without Bonferroni correction, for reasons details in Epskamp, Borsboom, & Fried 2017
plot(nonparam_boot, "strength") + theme(axis.text = element_text(size = 15), legend.text = element_text(size = 15))
print(plot(nonparam_boot, "closeness")) + theme(axis.text = element_text(size = 15), legend.text = element_text(size = 15))
print(plot(nonparam_boot, "betweenness")) + theme(axis.text = element_text(size = 15), legend.text = element_text(size = 15))
```

```{r}
strength_comparison_plot <- print(plot(nonparam_boot, "strength") + theme(text = element_text(size = 15)))
closeness_comparison_plot <- print(plot(nonparam_boot, "closeness")+ theme(text = element_text(size = 15)))
betweenness_comparison_plot <- print(plot(nonparam_boot, "betweenness")+ theme(text = element_text(size = 15)))
#ggsave('images/strength_comparison_plot.png', plot = strength_comparison_plot, width = 8, height = 5)
#ggsave('images/closeness_comparison_plot.png', plot = closeness_comparison_plot, width = 8, height = 5)
#ggsave('images/betweenness_comparison_plot.png', plot = betweenness_comparison_plot, width = 8, height = 5)
```

### Analyze both edge weight and node centrality stability using nodedrop bootstrapping

```{r}
# Conduct node-drop bootstrapping
if (!file.exists("data/nodedrop_boot.RDS")){
  nodedrop_boot <- bootnet(symptom_data, nBoots=2500, type="node", statistics=c('edge','strength','closeness','betweenness'), fun='network_estimation_fun')
  saveRDS(nodedrop_boot, file="data/nodedrop_boot.RDS")}
nodedrop_boot <- readRDS("data/nodedrop_boot.RDS")
```

```{r}
plot(nodedrop_boot, statistics=c('edge'))
plot(nodedrop_boot, statistics=c('strength','closeness','betweenness'))
```

```{r, fig.width=3.6, fig.height=3.5}
plot(nodedrop_boot, statistics=c('edge')) + theme_minimal() + xlab("% sampled nodes") + ylab("correlation with original") + theme(axis.text.y =element_text(size=20),axis.text.x=element_text(size=20),axis.title =element_text(size=20), legend.text = element_text(size=30), legend.position = "top")
```

```{r, fig.width=3.6, fig.height=3.5}
tolower_centrality_colors <- centrality_colors
names(tolower_centrality_colors) <- tolower(names(tolower_centrality_colors))
tobridge_centrality_colors <- centrality_colors
names(tobridge_centrality_colors) <- sapply(names(tobridge_centrality_colors), function(x) paste0("bridge",x))

plot(nodedrop_boot, statistics=c('strength','closeness','betweenness')) + scale_color_manual(values=tolower_centrality_colors, labels=names(centrality_colors)) + scale_fill_manual(values=tolower_centrality_colors, labels=names(centrality_colors)) + theme_minimal() + xlab("% sampled nodes") + ylab("correlation with original") + theme(axis.text.y =element_text(size=20),axis.text.x=element_text(size=20),axis.title =element_text(size=20), legend.text = element_text(size=20), legend.position = "top") + guides(fill=guide_legend(title=""), color=guide_legend(title=""))
```

```{r}
plot(nodedrop_boot, statistics=c('strength'), perNode=TRUE) + xlab("% sampled nodes") + theme(strip.text=element_text(size=15), axis.title.x =element_text(size=15), axis.text.y=element_text(size=11), axis.text.x=element_text(size=11))
plot(nodedrop_boot, statistics=c('closeness'), perNode=TRUE) + xlab("% sampled nodes") + theme(strip.text=element_text(size=15), axis.title.x =element_text(size=15), axis.text.y=element_text(size=11), axis.text.x=element_text(size=11))
plot(nodedrop_boot, statistics=c('betweenness'), perNode=TRUE) + xlab("% sampled nodes") + theme(strip.text=element_text(size=15), axis.title.x =element_text(size=15), axis.text.y=element_text(size=11), axis.text.x=element_text(size=11))
```
